A Hu-Washizu variational consistent meshfree thin shell formulation with naturally accommodating essential boundary conditions

### 2024-02-13 examples.tex
turn light

\section{Numerical examples}
In this section, several examples are carried out to verify proposed method, which employs the consistent reproducing kernel gradient smoothing integration scheme (RKGSI) and the non-consistent Gauss integration scheme (GI) with penalty method, Nitsche's method and the proposed Hu-Washizu formulation (HW) to enforce the essential boundary conditions. A normalized support size of 2.5 is used for all methods to ensure the requirement of quadratic base meshfree approximation. To eliminate the influence of integration, the Gauss integration scheme use 6 Gauss points for domain integration and 3 points for boundary integration, and such that the number of integration points are identical between Gauss scheme and RKGSI scheme. The error estimates of displacement namely $L_2$-Error and energy namely $H_e$-Error is used here:
\begin{equation}
\begin{split}
L_2\text{-Error} &= \frac{\sqrt{\int_\Omega(\boldsymbol v - \boldsymbol v^h) \cdot (\boldsymbol v - \boldsymbol v^h)d\Omega}}{\sqrt{\boldsymbol v \cdot \boldsymbol v}} \\
H_e\text{-Error} &= \frac{\sqrt{\int_\Omega \left ((\varepsilon_{\alpha\beta} - \varepsilon_{\alpha\beta}^h)(N^{\alpha\beta} - N^{\alpha\beta h}) + \int_\Omega(\kappa_{\alpha\beta}-\kappa_{\alpha\beta}^h)(M^{\alpha\beta}-M^{\alpha\beta h}) \right )d\Omega}}{\sqrt{\int_\Omega(\varepsilon_{\alpha\beta}N^{\alpha\beta} + \kappa_{\alpha\beta}M^{\alpha\beta})d\Omega}}
\end{split}
\end{equation}

>>>>

\section{Numerical examples}\label{examples}
In this section, several examples are carried out to verify proposed method, which employs the consistent reproducing kernel gradient smoothing integration scheme (RKGSI) and the non-consistent Gauss integration scheme (GI) with penalty method, Nitsche's method and the proposed Hu-Washizu formulation (HW) to enforce the essential boundary conditions. A normalized support size of 2.5 is used for all methods to ensure the requirement of quadratic base meshfree approximation. To eliminate the influence of integration, the Gauss integration scheme use 6 Gauss points for domain integration and 3 points for boundary integration, and such that the number of integration points are identical between Gauss scheme and RKGSI scheme. The error estimates of displacement namely $L_2$-Error and energy namely $H_e$-Error is used here:
\begin{equation}
\begin{split}
L_2\text{-Error} &= \frac{\sqrt{\int_\Omega(\boldsymbol v - \boldsymbol v^h) \cdot (\boldsymbol v - \boldsymbol v^h)d\Omega}}{\sqrt{\boldsymbol v \cdot \boldsymbol v}} \\
H_e\text{-Error} &= \frac{\sqrt{\int_\Omega \left ((\varepsilon_{\alpha\beta} - \varepsilon_{\alpha\beta}^h)(N^{\alpha\beta} - N^{\alpha\beta h}) + \int_\Omega(\kappa_{\alpha\beta}-\kappa_{\alpha\beta}^h)(M^{\alpha\beta}-M^{\alpha\beta h}) \right )d\Omega}}{\sqrt{\int_\Omega(\varepsilon_{\alpha\beta}N^{\alpha\beta} + \kappa_{\alpha\beta}M^{\alpha\beta})d\Omega}}
\end{split}
\end{equation}

### 2024-02-13 frontmatter.tex
turn light

\begin{frontmatter}
    \title{A principle meshfree formulation for thin shells with naturally consistent enforcement of essential boundary conditions}
    \author[1]{Junchao Wu\corref{cor1}}
    \ead{jcwu@hqu.edu.cn}
    \author[2]{Yangtao Xu}
    \author[1]{Bin Xu}
    \author[3]{Syed Humayun Basha}

@@

\begin{abstract}
Thin shell problems ignore the shear deformations and this leads to a requirement of C1 continuous approximations. Meshfree methods equipped with high order smoothed shape functions is suitable for thin shell analysis, since the high order shape function can also suppress the membrane locking in thin shell problems. However, meshfree shape function always perform a natural rational property, this is a big challenge to meet integration consistency for traditional Gauss integration rule within Galerkin weak form, while integration consistency serves a key role in accuracy of Galerkin meshfree methods. In this work, we proposed a reproducing kernel gradient smoothing integration (RKGSI) algorithm for thin shell problems, while the first and second order smoothed gradients are constructed based upon reproducing kernel smoothing gradient framework, with the aid of this framework, the integration consistency becomes a natural property by a replacement between smoothed gradients and traditional gradients of shape functions in Galerkin weak form. The order of basis functions used in smoothed gradient is determined by ensuring the optimal order of error convergence respected to energy norm. The traditional costly second order gradients are totally eliminated in RKGSI formulation. To further increase the efficiency of proposed method, a set of integration schemes are developed for consistent assembly of stiffness matrix, force vector and smoothed gradients, where the number of integration points, which accompanied with calculation of traditional shape functions and their first order gradients, are minimized by a global point of view. It is evident that the smoothed gradients meets the reproducing consistency of gradients that can ensure the optimal convergence property. The numerical examples demonstrate the efficacy and efficiency of proposed method, while the RKGSI performs a comparable result in energy error with interpolation by meshfree approximations.
\end{abstract}
\begin{keyword}
Meshfree \sep Thin shell \sep Hu-Washizu variational principle \sep Reproducing kernel gradient smoothing \sep Essential boundary condition
\end{keyword}
\end{frontmatter}

>>>>

\begin{frontmatter}
    \title{A Hu-Washizu variational consistent meshfree thin shell formulation with naturally accommodating essential boundary conditions}
    \author[1]{Junchao Wu\corref{cor1}}
    \ead{jcwu@hqu.edu.cn}
    \author[2]{Yangtao Xu}
    \author[1]{Bin Xu}
    \author[3]{Syed Humayun Basha}

@@

\begin{abstract}
A Hu-Washizu principle based variational consistent meshfree formulation with naturally enforcement of essential boundary conditions is proposed for thin shell analysis. In this approach, a mixed formulation of displacements, strains and stresses within the framework of Hu-Washizu variational principle is employed, where the displacements are discretized by meshfree shape functions, the strains and stresses are expressed as the smoothed gradients and covariant smoothed gradients which meet the first two order integration constraint and have the quasi- varational consistency. 
The numerical examples demonstrate the efficacy and efficiency of proposed method, while the RKGSI performs a comparable result in energy error with interpolation by meshfree approximations.
\end{abstract}
\begin{keyword}
Meshfree \sep Thin shell \sep Hu-Washizu variational principle \sep Reproducing kernel gradient smoothing \sep Essential boundary condition
\end{keyword}
\end{frontmatter}

### 2024-02-13 introduction.tex
turn light

\section{Introduction}
Thin shell is one of the most frequently used structure in engineering practice, where the thickness of this kind structure is often much smaller than its radius. With the Kirchhoff-Love hypothesis [1–3], the transverse shear deformation is eliminated in thin shell analysis, such that at least C1 continuous shape functions are required within Galerkin methods. In static and dynamic simulation of structure, the conventional finite element methods [1,2] are one of the most popular approximation scheme, however the construction of C1 continuity is still a big challenge for cell-based finite element methods. In last three decades, the meshfree methods [1–3] equipped high order smoothed shape functions have attracted significant research attention, while the meshfree shape functions are established based upon a set of scattered nodes and the high order continuity of shape functions is easily fulfilled even with low order basis function. For thin shell analysis, this high order meshfree approximations can also alleviate the membrane locking caused by the mismatched approximation order of membrane strain and bending strain [1]. Moreover, in general, the nodal-based meshfree approximations can release the burden of mesh distortion and have the flexibility of local refinement. Due to these advantages, a wide variety meshfree methods are proposed and have been applied to many scientific or engineering fields. Among of them, moving least squares (MLS) and reproducing kernel (RK) meshfree approximations built their shape functions by enforcing the so-call consistency conditions, where the consistency conditions require that the corresponding approximations should exactly reproduce every functions spanned by basis functions, and this conditions usually serve as a basic requirement for the error convergence of resolved Galerkin solutions [1]. However, the high order smoothed meshfree shape functions accompany with the severely overlapping supports, which leads to a misalignment between numerical integration domains and supports of shape functions. As a result, the meshfree shape functions usually exhibit a piecewise rational nature in each integration domains, and it brings a serious difficulty to the accurate numerical integration in Galerkin weak forms [1].

>>>>

\section{Introduction}\label{introduction}
Thin shell structure follows the Kirchhoff hypothesis that neglects the shear deformation \cite{donnell1976}, which requires the approximation should have at least $C^1$ continuity in Galerkin formulations.
The traditional finite element methods usually only has $C^0$ continuous shape functions, and it more prefers Mindlin thick shear theory, hybrid and mixed models in simulation of shell structure \cite{hughes2000}.
In last three decades, the meshfree methods \cite{belytschko1994, liu1995, chen2017a} equipped high order smoothed shape functions have attracted significant research attention, while the meshfree shape functions are established based upon a set of scattered nodes and the high order continuity of shape functions is easily fulfilled even with low order basis function.
For thin shell analysis, this high order meshfree approximations can also alleviate the membrane locking caused by the mismatched approximation order of membrane strain and bending strain \cite{krysl1996}.
Moreover, in general, the nodal-based meshfree approximations can release the burden of mesh distortion and have the flexibility of local refinement.
Due to these advantages, a wide variety meshfree methods are proposed and have been applied to many scientific or engineering fields. 
However, the high order smoothed meshfree shape functions accompany with the enlarged and overlapping supports, which may also leads to many issues for shape functions. One is the loss of Kronecker delta property \cite{fernandez-mendez2004}, which leads to that the essential boundary conditions cannot be enforced directly like finite element methods.  
Another issue is that the variational consistency or said integration constraint cannot be satisfied, which is caused by the misalignment between numerical integration domains and supports of shape functions, and the shape functions exhibit a piecewise rational nature in each integration domains.
Variational consistency is of importance to the solution accuracy in Galerkin formulations \cite{wu2021}.

To directly enforce the essential boundary for Galerkin meshfree methods, several approaches have been proposed for the recovery of shape functions' Kronecker property. For examples, interpolation element-free method \cite{liu2019a}, mixed transformation method \cite{chen2000a}, boundary singular kernel method \cite{chen2000a} etc.
However, these methods are not based on a variational setting, and cannot guarantee the variational consistency, enforcing accuracy may be worse on where there is no meshfree node. In contrast, enforcing the essential boundary conditions by a variational approach are more preferred for Galerkin meshfree methods. 
Belytschko et al.  \cite{belytschko1994, krysl1996} firstly introduced the variational consistent Lagrange multiplier method to Galerkin meshfree method, in which the extra degrees of freedom should be employed for discretion of Lagrange multiplier. And this method has been extended to geometrically nonlinear thin shells by Ivannikov et al. \cite{ivannikov2014a}. 
To eliminate the extra degrees of freedom, Lu et al. \cite{lu1994} represented the Lagrange multiplier by corresponding tractions and proposed the modified variational essential boundary enforcement method. However, the coercivity of this approach is not always ensured and potentially reduces the accuracy. 
Zhu and Atluri \cite{zhu1998} pioneered the penalty method for meshfree method, making it straightforward approach for enforcing essential boundary conditions via Galerkin weak form. However, penalty method suffers from a lack of variational consistency, and requires the experimental artificial parameters, whose optimal value is hard to be determined.
Fernández-Méndez and Huerta \cite{fernandez-mendez2004} used the Nitsche's method in meshfree formulation for imposing essential boundary conditions. This method can be viewed as a hybrid of modified variational method and penalty method, since its consistent term that ensure variational consistency generated by modified variational method, and the penalty method is employed as stabilized term to recovery the coercivity. Skatulla and Sansour \cite{skatulla2008} further extended Nitsche's method for thin shell analysis and proposed an iteration algorithm to determine artificial parameters at each integration points.

To address the issue of numerical integration, a serial of consistent integration scheme has been developed for Galerkin meshfree methods. For instance, stabilized conforming nodal integration \cite{chen2001}, variational consistent integration \cite{chen2013}, quadratic consistent integration \cite{duan2012}, reproducing kernel gradient smoothing integration \cite{wang2019d}, consistent projection integration \cite{wang2023} etc.
The most consistent integration scheme is established by assumed strain approach, while the costly higher order derivatives of traditional meshfree shape functions are replaced by smoothed gradient, and show a high efficiency. 
Moreover, in order to achieve the global variational consistency, a consistent essential boundary condition enforcement should cooperate with the consistent integration scheme. 
The pair of consistent integration scheme and Nitsche's method for the treatment of essential boundary conditions shows a good performance, since it no needs the extra degrees of freedom and can fulfilled the coercivity. However,in Nitsche's method, the artificial parameters still exist in stabilized term and the costly higher order derivatives should be recalled, especially for thin plate and thin shell problems \cite{benzaken2021}.
Recently, Wu et al \cite{wu2022b, wu2023} proposed a efficient and stabilized essential boundary condition enforcement based upon the Hellinger-Reissner (HR) variational principle, where the reproducing kernel gradient smoothing integration is recast by a mixed formulation in Hellinger-Reissner weak form. The terms for enforcing essential boundary conditions is mostly identical with Nitsche's method, both have consistent term and stabilized term. Nevertheless, the stabilized term of this method naturally exist in Hellinger-Reissner weak form and no longer needs the artificial parameters, even for essential boundary enforcement, total of the higher order derivatives are represented by smoothed gradients and their derivatives.

In this study, an efficient and stabilized variational consistent meshfree method with naturally enforcing the essential boundary conditions is developed for thin shell structure. Follow the ideas of Hellinger-Reissner principle base consistent meshfree method, the Hu-Washizu variational principle of complementary energy \cite{dah-wei1985} with variables of displacement, strains and stresses is employed, where the displacement is approximated by conventional meshfree shape functions, and the strains and stresses are expressed by the smoothed gradients or covariant smoothed gradients with covariant bases. It should be noted that the smoothed gradients inherently embed the first two order integration constraints, however, due to the non-polynomial property of stresses, the fulfillment of these integration constraint only can get a quasi-satisfaction of variational consistency. All of the essential boundary conditions about displacements and rotations are considered in Hu-Washizu weak form, and present a Nitsche-like formalism but without any artificial parameters. Comparing with Nitsche's method, the costly higher order derivatives are replaced by conventional reproducing smoothed gradients and its direct derivatives. Taking the advantages of reproducing kernel gradient smoothing framework, the smoothed gradients shows a better performance on efficiency than conventional derivatives of shape functions, which improves the computational efficiency of meshfree formulation.

The remainder of this paper is organized as follows.
Section \ref{Kinematics} briefly describes the kinematics of thin shell structure and the corresponding Hu-Washizu principle weak form.
Subsequently, the mixed formulation regarding the displacements, strains and stresses in accordance with Hu-Washizu weak form is presented in Section \ref{mixed}.
Section \ref{boundary} derives the discrete equilibrium equations with the naturally accommodation of essential, and compares them with those of Nitsche's method.
The efficacy of the proposed Hu-Washizu meshfree thin shell formulation is validated by numerical results in Section \ref{examples}.
Concluding remarks are finally drawn in Section \ref{conclusion}.
